Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-07-05 survey/survey.tex
tweak some level sentences. None were wrong.

Starting with the simplest constructions, the first two constructs in
Fig.~\ref{fige} are the level-2 analogs of $+$ and $*$, which respectively
model combinator-level choice composition and parallel composition (of
equivalences).  These allow us to ``build up'' larger equivalences from smaller
ones.  The next two express that both of these composition operators distribute
over sequential composition $\odot$ (and vice versa).

@@

The constructs in Fig.~\ref{figj} capture the informal idea that all
the different ways of associating programs are equivalent. The first
says that sequential composition itself ($\odot$) is associative.
The next $4$ capture how
the $\oplus$ and $\otimes$ combinators ``commute'' with re-association.
In other words, it expresses the type-level associativity of $+$ is
properly reflected by the properties of $\oplus$.
The last two equivalences show how composition of associativity combinators
interact together.

@@

The constructs in Fig.~\ref{figi} are the basic coherence for
$\dist$, $\distl$, $\factor$ and $\factorl$: the type-level distribution
and factoring has to commute with the combinator-level $\oplus$ and $\otimes$.

@@

The first two equivalences in Fig.~\ref{figf} reflect the basic
coherence between type-level swapping and and the level-2 combinator
actions. The next four arise because of interactions between
(additive and multiplicative) level-1 associativity and swapping.
In other words, they arise as critical pairs.  For example,
the first expresses that the two ways of going from
$\left(A \oplus B\right) \oplus C$ to $B \oplus \left(C \oplus A\right)$
are equivalent, with the second saying that the reverse (i.e.
the results of applying $!$\,) also gives equivalent programs.
The last two say the same but for the multiplicative structure.

@@

\newcommand{\transLR}{\AgdaInductiveConstructor{trans⇔}}
If all expressible transformations were single-step only, this
would hardly justify calling this an ``operational semantics'',
as we would hardly have a programming language. However, this level
of $\Pi$ has combinators as well: two are in
Fig.~\ref{fige} and two are in Fig.~\ref{figh}. The most interesting
one is ``sequential composition'', which is the middle one at the 
bottom of Fig.~\ref{figh}. Since $\Leftrightarrow$ represents
an equivalence, sequential composition in this context is the
same as transitivity of equivalences, as thus we have chosen to
name this \transLR. When evaluating \transLR, we could cheat:
we know that the eventual answer must be, and we could just 
return that. But this is not operational in any real sense, as
that skips over the intermediate steps. We would like to be able
to ``trace'' the rewrite. Thus the evaluation of
$\transLR r_0 r_1$ where $r_0 : c_0 \leftrightarrow c_1$ and
$r_1 : c_1 \leftrightarrow c_2$
should apply \evalone to both $r_0$ and $r_1$. Furthermore,
after applying $r_0$, we should be able to witness that the result
is indeed $c_1$, so that we may continue. This last requirement
forces us to define a new function, mutually recursively with
\evalone, for this task:
\begin{flalign*}
\mathit{exact}\,:\,\ \left\{ t_1\, t_2\,:\,U\right\}
  \left\{ c_1\, c_2\,:\, t_1\,\leftrightarrow\,t_2\right\}
  (ce\,\,c_1 \Leftrightarrow\,c_2) \rightarrow \evalone\, ce \equiv c_2 \\
\end{flalign*}
\noindent If we are careful in our construction of \evalone, the
definition of $\mathit{exact}$ is quite straightforward, i.e.
almost all cases are \AgdaInductiveConstructor{refl}.

>>>>

Starting with the simplest constructions, the first two constructs in
Fig.~\ref{fige} are the level-2 analogs of $+$ and $*$, which respectively
model level-1 choice composition and parallel composition (of
equivalences).  These allow us to ``build up'' larger equivalences from smaller
ones.  The next two express that both of these composition operators distribute
over sequential composition $\odot$ (and vice versa).

@@

The constructs in Fig.~\ref{figj} capture the informal idea that all
the different ways of associating programs are equivalent. The first
says that sequential composition itself ($\odot$) is associative.
The next $4$ capture how
the $\oplus$ and $\otimes$ combinators ``commute'' with re-association.
In other words, it expresses that the type-level associativity of $+$ is
properly reflected by the properties of $\oplus$.
The last two equivalences show how composition of associativity combinators
interact together.

@@

The constructs in Fig.~\ref{figi} are the basic coherence for
$\dist$, $\distl$, $\factor$ and $\factorl$: the type-level distribution
and factoring has to commute with the level-1 $\oplus$ and $\otimes$.

@@

The first two equivalences in Fig.~\ref{figf} reflect the basic
coherence between level-0 swapping and the level-1 combinator
actions. The next four arise because of interactions between
(additive and multiplicative) level-1 associativity and swapping.
In other words, they arise as critical pairs.  For example,
the first expresses that the two ways of going from
$\left(A \oplus B\right) \oplus C$ to $B \oplus \left(C \oplus A\right)$
are equivalent, with the second saying that the reverse (i.e.
the results of applying $!$\,) also gives equivalent programs.
The last two say the same but for the multiplicative structure.

@@

\newcommand{\transLR}{\AgdaInductiveConstructor{trans⇔}}
If all expressible transformations were single-step only, this
would hardly justify calling this an ``operational semantics'',
as we would hardly have a programming language. However, level-2
of $\Pi$ has combinators as well: two are in
Fig.~\ref{fige} and two are in Fig.~\ref{figh}. The most interesting
one is ``sequential composition'', which is the middle one at the 
bottom of Fig.~\ref{figh}. Since $\Leftrightarrow$ represents
an equivalence, sequential composition in this context is the
same as transitivity of equivalences, as thus we have chosen to
name this \transLR. When evaluating \transLR, we could cheat:
we know that the eventual answer must be, and we could just 
return that. But this is not operational in any real sense, as
that skips over the intermediate steps. We would like to be able
to ``trace'' the rewrite. Thus the evaluation of
$\transLR r_0 r_1$ where $r_0 : c_0 \leftrightarrow c_1$ and
$r_1 : c_1 \leftrightarrow c_2$
should apply \evalone to both $r_0$ and $r_1$. Furthermore,
after applying $r_0$, we should be able to witness that the result
is indeed $c_1$, so that we may continue. This last requirement
forces us to define a new function, mutually recursively with
\evalone, for this task:
\begin{flalign*}
\mathit{exact}\,:\,\ \left\{ t_1\, t_2\,:\,U\right\}
  \left\{ c_1\, c_2\,:\, t_1\,\leftrightarrow\,t_2\right\}
  (ce\,\,c_1 \Leftrightarrow\,c_2) \rightarrow \evalone\, ce \equiv c_2 \\
\end{flalign*}
\noindent If we are careful in our construction of \evalone, the
definition of $\mathit{exact}$ is quite straightforward, i.e.
almost all cases are \AgdaInductiveConstructor{refl}.
